---
output: 
  govdown::govdown_document:
    font: "sans-serif"
    favicon: "custom"
    logo: "images/uk_white_on_transparent.svg"
    logo_url: "https://co-analysis.github.io/csmapping/cs_lad.html"
    logo_text: ""
    page_title: "UK Civil Servants by local authority (2018)"
    title: "Mapping civil servants"
    phase: alpha
    feedback_url: "https://www.github.com/co-analysis/csmapping/issues"
    google_analytics: "UA-139844526-1"
    
---

```{r setup, include=FALSE}

library(tidyverse)
library(geojsonio)
library(sp)
library(leaflet)
library(DT)
library(knitr)
library(kableExtra)

cs18_lad <- read_csv("https://www.nomisweb.co.uk/api/v01/dataset/NM_129_1.data.csv?geography=1820327937...1820328307&date=latest&department=0&national_identity=0&gender=0&full_part_time=0&ethnic_group=0&disability_status=0&wage_band=0&civil_service_grade=0&age_band=0&measures=20100")

spdf_lad <- geojson_read("https://opendata.arcgis.com/datasets/bbb0e58b0be64cc1a1460aa69e33678f_0.geojson", what = "sp")

lad_dt <- cs18_lad %>%
  select(geo_id = GEOGRAPHY_CODE, name = GEOGRAPHY_NAME, value = OBS_VALUE) %>%
  mutate(pc = formattable::percent(
    value/430080),
    value = formattable::comma(value, digits = 0))

lad_leaf <- sp::merge(spdf_lad, lad_dt, by.x = "lad19cd", by.y = "geo_id")

bincol <- colorBin("Purples",
                   lad_dt$value,
                   bins = c(0, 500, 1000, 2500, 5000, 7000, 40000, 50000),
                   pretty = FALSE,
                   na.color = "#eeeeee")

```

# UK Civil Servants by local authority (2018)

::: {.lead-para}
This page provides analysis of UK civil servants by local authority (district and unitary authorities) in England, Scotland and Wales, as at 31 March 2018.
:::

The map shows the number of UK civil servants working in 369 local authorities in England, Scotland and Wales. There are 3,540 civil servants working in Northern Ireland for UK Civil Service organisations, where data is not available by local authority. There are also a further 7,440 UK civil servants who work overseas or whose working location is not known.

The table shows the data for all 369 local authorities in the map, in addition to the number of civil servants in that local authority it also shows this number as a percentage of all UK Civil Servants.

Figures are rounded to the nearest 10, data has been supressed for two local authorities where figures are less than five (including zero).

```{r map, echo=FALSE}

leaflet(lad_leaf, width = "100%", height = 500) %>%
  addProviderTiles(providers$CartoDB.PositronNoLabels) %>%
  addMapPane("dt", zIndex = 410) %>%
  addMapPane("labs", zIndex = 420) %>%
  addPolygons(color = "#aaaaaa",
              weight = 1,
              fillColor = ~bincol(value),
              popup = ~paste(lad19nm, value, sep = ": "),
              fillOpacity = 0.8,
              options = pathOptions(pane = "dt")) %>%
  addProviderTiles(providers$CartoDB.PositronOnlyLabels, 
                   options = providerTileOptions(pane = "labs")) %>%
  addLegend(position = "bottomleft", pal = bincol, values = ~lad_dt$value, 
            title = "Number of civil servants", opacity = 0.8)

```


```{r table, echo=FALSE}

lad_dt2 <- lad_dt %>%
  select(name, value, pc) %>%
  arrange(desc(value)) %>%
  drop_na(value)

datatable(lad_dt2,
          colnames = c("Local authority" = "name",
                       "Number" = "value",
                       "Percent of Total" = "pc"),
          options =
            list(pageLength = 15,
                 initComplete = JS("function(settings, json) {",
                                   "$('.datatables').css({'font-family': 'sans-serif'});",
                                   "}"))) %>%
  formatRound("Number", digits = 0) %>%
  formatPercentage("Percent of Total", 1)

```

---

#### Who are UK civil servants?
UK civil servants work for organisations in the [UK Civil Service](https://www.gov.uk/civil-service), supporting the UK Government, the Scottish Government and the Welsh Government (e.g. government departments, executive agencies or Crown Non-Departmental Public Bodies).

#### What about civil servants in Northern Ireland?
There are 3,540 UK civil servants who work in Northern Ireland. However, most civil servants in Northern Ireland work for the [Northern Ireland Civil Service (NICS)](https://www.finance-ni.gov.uk/topics/working-northern-ireland-civil-service), a separate entity from the UK Civil Service. Statistics about the NICS are available from the [Northern Ireland Statistics and Research Agency (NISRA)](https://www.nisra.gov.uk/statistics/government/ni-civil-service-human-resource-statistics).

#### About this analysis
The data for the map and table comes from the [Annual Civil Service Employment Survey (ACSES)](https://www.gov.uk/government/collections/civil-service-statistics), provided via the [NOMIS API](https://www.nomisweb.co.uk/), the geographic shape files are from the ONS's [OpenGeography Portal](http://geoportal.statistics.gov.uk). The local authority a civil servant works in is defined in reference to the postcode of their main working location recorded in Civil Service HR systems. This document is produced in [RMarkdown](http://rmarkdown.rstudio.com) using [`leaflet`](https://rstudio.github.io/leaflet/) and [`DT`](https://rstudio.github.io/DT/); it has been rendered into HTML using [`govdown`](https://ukgovdatascience.github.io/govdown/).
