---
output: 
  govdown::govdown_document:
    font: "sans-serif"
    favicon: "custom"
    logo: "images/uk_white_on_transparent.svg"
    logo_url: "https://co-analysis.github.io/csmapping/cs_lad_2019.html"
    logo_text: ""
    page_title: "UK civil servants by postcode (2019)"
    title: "Mapping the UK Civil Service"
    phase: alpha
    feedback_url: "https://www.github.com/co-analysis/csmapping/issues"
    google_analytics: "UA-139844526-1"
    
---

```{r setup, include=FALSE}

# load packages
library(tidyverse)
library(geojsonio)
library(sf)
library(leaflet)
library(DT)
library(knitr)
library(crosstalk)

# urls for downloading assets
urls <- list(
  cs19pc = "https://assets.publishing.service.gov.uk/government/uploads/system/uploads/attachment_data/file/837270/Civil-servants-by-postcode-organisation-grade-and-leaving-cause__1__-_Staff_in_Post.csv",
  ons_lad = "https://opendata.arcgis.com/datasets/c3ddcd23a15c4d7985d8b36f1344b1db_0.geojson"
)

# read in civil service stats and filter to table 15
cs19pc <- read_csv(urls$cs19pc, 
                   skip = 8, 
                   col_names = c("organisation", "postcode", "aoaa", "eo", 
                                 "seoheo", "g6g7", "scs", "unknown", "total"), 
                   na = "..")

cs19pc_clean <- cs19pc %>%
  select(organisation, postcode, total) %>%
  drop_na(total) %>%
  arrange(organisation) %>%
  mutate(organisation = as_factor(organisation))

onspd_slim <- vroom::vroom("onspd_slim_201908.csv")

ons_lad <- geojsonio::geojson_read(urls$ons_lad, parse = TRUE)

ons_lad_dt <- ons_lad$features$properties 

ons_lad_dt <- ons_lad_dt %>% select(local_authority = LAD19NM, oslaua = LAD19CD)

onspd_slim2 <- onspd_slim %>%
  select(pcds, oslaua, lat, long) %>%
  mutate(postcode = str_remove(pcds, "\\s")) %>%
  filter(postcode %in% cs19pc_clean$postcode)

cs19pc_merged <- cs19pc_clean %>%
  left_join(onspd_slim2, by = c("postcode" = "postcode")) %>%
  drop_na(oslaua) %>%
  left_join(ons_lad_dt) %>%
  arrange(desc(total))

cs19pc_sf <- st_as_sf(cs19pc_merged, coords = c("long", "lat"))

sd_cs19pc <- SharedData$new(cs19pc_sf)

```

# UK civil servants by postcode and organisation (2019)

::: {.lead-para}
This page provides a map and tables of UK civil service headcount by postcode and organisation, as at 31 March 2019.
:::

For full details about the locations that civil servants work in please see [Civil Servants by local authority](cs_lad_2019.html).

The map below expands on this analysis by looking at the distribution of civil servants by postcode and organisation. Each circle on the map represents a group of civil servants working for a government department or executive agency in a particular postcode. The size of the circle is proportionate to the headcount of civil servants working in that organisation and postcode.

:::{.govuk-grid-row}

:::{.govuk-grid-column-two-thirds}

``` {r map, echo=FALSE, warning=FALSE}
leaflet(sd_cs19pc, width = "100%", height = 600) %>%
  addProviderTiles(providers$CartoDB.PositronNoLabels) %>%
  addMapPane("dt", zIndex = 410) %>%
  addMapPane("labs", zIndex = 420) %>%
  addCircles(radius = ~sqrt(total)*100,
             popup = ~paste(organisation, pcds, formattable::comma(total, digits = 0), sep = "; "),
             weight = 2, fillColor = "#005abb", color = "#005abb", opacity = 0.6,
             options = pathOptions(pane = "dt")) %>%
  addProviderTiles(providers$CartoDB.PositronOnlyLabels, 
                   options = providerTileOptions(pane = "labs"))

```

:::

:::{.govuk-grid-column-one-third}

::: {.small-para}
You can select local authorities and/or organisations of interest. Note that due to technical limitations the map will not automatically zoom in to your selection.

**Select local authorities:**
```{r select_la, echo=FALSE, warning=FALSE}

fs_la <- filter_select(id = "select_la", label = "", sharedData = sd_cs19pc,
              group = ~local_authority)

fs_la_nobootstrap <- fs_la

attr(fs_la_nobootstrap, "html_dependencies") <- Filter(
  function(dep) {dep$name != "bootstrap"},
  attr(fs_la_nobootstrap, "html_dependencies")
)

fs_la_nobootstrap

```


**Select organisations:**
```{r select_org, echo=FALSE, warning=FALSE}

fs_org <- filter_select(id = "select_org", label = "", sharedData = sd_cs19pc,
              group = ~organisation)

fs_org_nobootstrap <- fs_org

attr(fs_org_nobootstrap, "html_dependencies") <- Filter(
  function(dep) {dep$name != "bootstrap"},
  attr(fs_org_nobootstrap, "html_dependencies")
)

fs_org_nobootstrap

```

:::

:::

:::
